---
title: "World distribution of Varroa mites mtDNA lineages"
author: "Maeva A. Techer, Ecology and Evolution OIST"
date: '2020-02-01'
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    depth: 3
    number_sections: no
    theme: yeti
    toc: yes
    toc_float: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document was written in R Markdown, and translated into html using the R package `knitr`. Press the buttons labelled **Code** to show the R code used to produce each interactive map created with `leaflet package`.

# Distribution of the honey bee Varroa mites haplogroups

**Background:** Varroa mites are specialist ectoparasites of _Apis_ honey bees and native from Asia. Four species have been described using morphometrics and mitochondrial markers. Besides, three mites had unresolved taxonomical status meaning that Varroa cryptic complex species is more diverse. Among them, _Varroa destructor_ and its sister species _V. jacobsoni_ have repeatedly and independently switch host from the Eastern honey bee _A. cerana_ to the varroa-naive Western honey bee _A. mellifera_. While _V. jacobsoni_ is for now restricted to the Guinean region following its jump in 2008, _V. destructor_ has quickly spread nearly worldwide since the 1960s. Only Australia with high biosecurity standard and remote islands such as Seychelles and Comoros archipelagos remain Varroa-free.

Two _V. destructor_ haplogroups based on the mtDNA _COX1_ partial gene, referred as the Korean K1 and the Japanese J1, were able to establish on _A. mellifera_. However, only K1 strain is reported cosmopolite and assumed to be more virulent and displacing the J1 strain in co-introduced areas. Although Varroa presence has been reported in many countries as a notifiable animal disease (OIE), strain or lineage identification is not always available and even more so over time.

To better understand the temporal dynamic of _V. destructor_ populations during the worldwide invasion, we visually reported the distribution of species and haplogroup/strains only confirmed by mtDNA _COX1_ sequencing or PCR-RFLP. Such an approach has previously provided a distribution map emphasizing the supposed parapatry trend between the two sister species _V. destructor_ and _V. jacobsoni_ found on their original host as well as the sympatry with _V. underwoodi_ (Techer et al. 2019). 


### Load R libraries 
```{r message=FALSE, warning=FALSE, results="hide"}
library(maps)
library(ggplot2)
library(tidyverse)
library(leaflet)
library(htmltools)
library(rgdal)
library(knitr)
library(kableExtra)

# Download the country borders layer
#download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="world_shape_file.zip")
#system("unzip world_shape_file.zip")
setwd("~/Documents/GitHub/varroa-mtDNA-world-distrib")
world_spdf=readOGR(dsn= getwd() , layer="TM_WORLD_BORDERS_SIMPL-0.3")
```

We reviewed 68 articles from 1995 to 2020 using either RAPD, mtDNA analysis (PCR-RFLP, sequencing) and/or nuclear microsatellites on Varroa mites or environmental honey DNA29. We collected distribution data about Varroa species identity, mtDNA haplogroup, date of sampling, geographical localization, honey bee host. For geographical localization, three cases occurred: a) exact coordinates were available, b) city or locality was available and c) no localization was available outside of the country level. In the second case, we approximated the geographical position as the center of the city/locality or placed it to the nearest Agricultural Institute or Academic Center/University as some past sampling was known to be carried in experimental research apiaries.. 
 
In addition, 485 mitochondrial sequences were downloaded from NCBI Genbank (last update on the 01 February 2020) for which for the same information if not included in the previous papers list. We obtained three mitogenomes, 387 partial _COX1_ sequences (length ranging from 188 to 1088bp), 36 partial _COX3_ sequences (ranging from 323 to 436 bp), 36 partial _ATP6_ sequences (ranging from 287 to 339 bp), 23 partial _CYTB_ sequences (ranging from 899 to 985 bp). All _COX1_ sequences were blasted and aligned to 44 reference mtDNA haplotypes (see Table S3)1,4,6,26,27,30–33. Despite highly variable length and inconsistent sequence overlapping due to the usage of variable primers and high sequences trimming, we expanded the known _COX1_ haplogroups . Following the nomenclature rules we found that _V. destructor_ is the most diverse with 31 haplogroups (including 22 K-like), followed by _V. jacobsoni_ with 19, five for _V. underwoodi_, for undetermined Varroa sp. and one for _V. rindereri_. We use these haplogroups to build an interactive distribution map of the Varroa mites on their honey bee hosts mikheyevlab.github.io/varroa-mtDNA-world-distrib/.

### Data collected from references to create map points
``` {r layersmap}
#Data with exact or approximate GPS coordinates obtained from references
coordvarroa <- read.csv("Data/mappingpoints_Varroareview05022020.csv", header = TRUE)
coord.vdac <- coordvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. cerana") 
coord.vjac <- coordvarroa %>% filter(Species == "V. jacobsoni") %>% filter(Host == "A. cerana")
```
  
In some cases, no other information than the country of sampling was reported. We build here a layer for which all countries in which one haplogroup was reported will be highlighted with the same color.  
  
_For example, I found no other mtDNA identification than Solignac et al. (2005) for Chile, so the country will be highlighted for V. destructor Korea presence._

### Data collected from references to create country polygons
```{r layercountry}
#Data with only the country name when no GPS data was available
paysvarroa <- read.csv("Data/mappingcountry_Varroareview05022020.csv", header = TRUE)

```
  
  
The following interactive map allows to plot the destribution of both invasive lineage for V. destructor on its novel host A. mellifera in regards with the recent expansion of its simili, *V. jacobsoni* in Papua New Guinea.  

It is possible to zoom in and out, and passing the cursor on one point will give basic information on the year and reporter. On the top right corner of the map, a control button allow to select specific points/country layers.  

Points and country are colored as following:  
<span style="color:red"> V. destructor **K lineage** </span>  
<span style="color:blue"> V. destructor **J lineage** </span>  
<span style="color:purple"> V. destructor **K and J lineages** coexist(ed) </span>  
<span style="color:green"> V. jacobsoni **Java lineage**</span>  
<span style="color:black"> Varroa sp. **Luzon lineage**</span>  

## Distribution of Varroa spp. mtDNA lineages on Apis mellifera
```{r vdamsubset}

### Subset the data for only Varroa destructor on the Western honey bee
coord.vdamK <- coordvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. mellifera") %>% filter(GROUP == "VD_K_type")

pays.vdamK <- paysvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. mellifera") %>% filter(GROUP == "VD_K_type")
loc.vdamK <- pays.vdamK$Location
map.vdamK <- subset(world_spdf,world_spdf@data$NAME %in% loc.vdamK)

coord.vdamJ <- coordvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. mellifera") %>% filter(GROUP == "VD_Japan_J1")

pays.vdamJ <- paysvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. mellifera") %>% filter(GROUP == "VD_Japan_J1")
loc.vdamJ <- pays.vdamJ$Location
map.vdamJ <- subset(world_spdf,world_spdf@data$NAME %in% loc.vdamJ)

coord.vdamKJ <- coordvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. mellifera") %>% filter(GROUP == "KoreaJapan")

pays.vdamKJ <- paysvarroa %>% filter(Species == "V. destructor") %>% filter(Host == "A. mellifera") %>% filter(GROUP == "KoreaJapan")
loc.vdamKJ <- pays.vdamKJ$Location
map.vdamKJ <- subset(world_spdf,world_spdf@data$NAME %in% loc.vdamKJ)

### Subset the data for only Varroa jacobsoni on the Western honey bee
coord.vjam <- coordvarroa %>% filter(Species == "V. jacobsoni") %>% filter(Host == "A. mellifera")

pays.vjam <- paysvarroa %>% filter(Species == "V. jacobsoni") %>% filter(Host == "A. mellifera")
loc.vjam <- pays.vjam$Location
map.vjam <- subset(world_spdf,world_spdf@data$NAME %in% loc.vjam)

### Subset the data for only Varroa sp. on the Western honey bee
coord.vspam <- coordvarroa %>% filter(Species == "Varroa sp.") %>% filter(Host == "A. mellifera")

pays.vspam <- paysvarroa %>% filter(Species == "Varroa sp.") %>% filter(Host == "A. mellifera")
loc.vspam <- pays.vspam$Location
map.vspam <- subset(world_spdf,world_spdf@data$NAME %in% loc.vspam)

# Prepare the map title
tag.map.title <- tags$style(HTML("
  .leaflet-control.map-title { 
    transform: translate(-50%,20%);
    text-align: center;
    padding-left: 10px; 
    padding-right: 10px; 
    background: rgba(255,255,255,0.75);
    font-weight: bold;
    font-size: 20px;
  }
"))

title <- tags$div(
  tag.map.title, HTML("Reported distribution of Varroa mite strains on A. mellifera")
)  

## Create the interactive map with leaflet
leaflet(coordvarroa) %>% 
  setView(lng = 0, lat = 0, zoom = 2) %>% 
  addTiles(group = "OSM (default)") %>%
  ## add the layer other than default we would like to use for background
  addProviderTiles(providers$Esri.WorldGrayCanvas, group = "Esri.WorldGrayCanvas") %>%
  addProviderTiles(providers$Stamen.TonerLines, group = "Stamen.TonerLines") %>%
  ## add the three layers coloring the countries in which K, J or both were described
  addPolygons(data = map.vdamK, fillColor = "red", stroke=FALSE, group = "Country with K strain") %>%
  addPolygons(data = map.vdamJ, fillColor = "blue", stroke=FALSE, group = "Country with J strain") %>%
  addPolygons(data = map.vdamKJ, fillColor = "#d9b7eb", stroke=FALSE, group = "Country co-infected") %>%
  ## adding Varroa jacosboni country layer
    addPolygons(data = map.vjam, fillColor = "black", stroke=FALSE, group = "Country with V. jacobsoni") %>%
    ## adding Varroa sp. country layer
    addPolygons(data = map.vspam, fillColor = "black", stroke=FALSE, group = "Country with Varroa sp.") %>%
  ## adding the three layers with GPS points for each jumping parasite strains
  addCircleMarkers(data = coord.vdamK, coord.vdamK$coord.Y, coord.vdamK$coord.X,
                   weight = 0.5,
                   col = "#FB0000", 
                   radius = 3.5, 
                   fillOpacity = 0.8, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor Korean lineage') %>%
  addCircleMarkers(data = coord.vdamJ, coord.vdamJ$coord.Y, coord.vdamJ$coord.X,
                   weight = 0.5,
                   col = "#068FF9", 
                   radius = 3.5, 
                   fillOpacity = 0.8, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor Japan lineage')
%>%
  addCircleMarkers(data = coord.vdamKJ, coord.vdamKJ$coord.Y, coord.vdamKJ$coord.X,
                   weight = 0.5,
                   col = "#9707E7", 
                   radius = 3.5, 
                   fillOpacity = 0.8, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'Co-infection by K & J') %>%
  ## adding Varro jacobsoni layer to show parapatry
    addCircleMarkers(data = coord.vjam, coord.vjam$coord.Y, coord.vjam$coord.X,
                   weight = 0.5,
                   col = "#black", 
                   radius = 3.5, 
                   fillOpacity = 0.8, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Java lineage') %>%
    ## adding Varroa sp. layer to show allopatry
    addCircleMarkers(data = coord.vspam, coord.vspam$coord.Y, coord.vspam$coord.X,
                   weight = 0.5,
                   col = "black", 
                   radius = 3.5, 
                   fillOpacity = 0.8, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'Varroa sp.') %>%
  ## adding the control button to remove or add layers of points or country
  addLayersControl(position = "topright",
    baseGroups = c("OSM (default)", "Positron NoLabels"),
    overlayGroups = c("V. destructor Korean lineage", 
                      "Country with K strain", 
                      "V. destructor Japan lineage", 
                      "Country with J strain", 
                      "Co-infection by K & J", 
                      "Country co-infected", 
                      "V. jacobsoni Java lineage", 
                      "Country with V. jacobsoni", 
                      "Varroa sp.", "Country with Varroa sp."),
    options = layersControlOptions(collapsed = TRUE)
  )  %>%
  ## adding a title for the map
  addControl(title, position = "bottomright", className="map-title") %>% 
  ## show the positron background prerably to the OSM layer
  showGroup("Positron NoLabels")
```
  
This map now focus on the diversity within *V. destructor* on its native host *A. cerana*.  
  
## Distribution of Varroa destructor mtDNA lineages on Apis cerana
```{r destructor}

vdac.K1 <- coord.vdac[coord.vdac$GROUP =="VD_K_type",]
vdac.J1 <- coord.vdac[coord.vdac$GROUP =="VD_Japan_J1",]
vdac.C1 <- coord.vdac[coord.vdac$GROUP =="VD_China_C1",]
vdac.C2 <- coord.vdac[coord.vdac$GROUP =="VD_China_C2",]
vdac.C3 <- coord.vdac[coord.vdac$GROUP =="VD_China_C3",]
vdac.C4 <- coord.vdac[coord.vdac$GROUP =="VD_China_C4",]
vdac.Nep <- coord.vdac[coord.vdac$GROUP =="VD_Nepal_1",]
vdac.SI <- coord.vdac[coord.vdac$GROUP =="VD_South_India_1",]
vdac.SRI <- coord.vdac[coord.vdac$GROUP =="VD_Sri_Lanka_1",]
vdac.V1 <- coord.vdac[coord.vdac$GROUP =="VD_Viet_Nam_V1",]

# Prepare the map title
tag.map.title2 <- tags$style(HTML("
  .leaflet-control.map-title2 { 
    transform: translate(-50%,20%);
    text-align: center;
    padding-left: 10px; 
    padding-right: 10px; 
    background: rgba(255,255,255,0.75);
    font-weight: bold;
    font-size: 20px;
  }
"))

title <- tags$div(
  tag.map.title2, HTML("Reported distribution of V. destructor strains on A. cerana")
)  

leaflet(coord.vdac) %>% 
  addTiles(group = "OSM (default)") %>%
  ## add the layer other than default we would like to use for background
  addProviderTiles(providers$CartoDB.PositronNoLabels, group = "Positron NoLabels") %>%
  ## adding each lineage group as a layer for V. destructor on cerana
    ## adding distribution of V. jacobsoni on original host A. cerana
  addCircleMarkers(data = coord.vjac, coord.vjac$coord.Y, coord.vjac$coord.X,
                   weight = 0.5,
                   col = "black", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni')%>%
  
  addCircleMarkers(data = vdac.K1, vdac.K1$coord.Y, vdac.K1$coord.X,
                   weight = 0.5,
                   col = "#FB0000", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor K1') %>%
  addCircleMarkers(data = vdac.J1, vdac.J1$coord.Y, vdac.J1$coord.X,
                   weight = 0.5,
                   col = "#0066ff", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor J1') %>%
  addCircleMarkers(data = vdac.C1, vdac.C1$coord.Y, vdac.C1$coord.X,
                   weight = 0.5,
                   col = "#ff9933", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor China C1') %>%
  addCircleMarkers(data = vdac.C2, vdac.C2$coord.Y, vdac.C2$coord.X,
                   weight = 0.5,
                   col = "#b800e6", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor China C2') %>%
  addCircleMarkers(data = vdac.C3, vdac.C3$coord.Y, vdac.C3$coord.X,
                   weight = 0.5,
                   col = "#ff4dff", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor China C3')  %>%
  addCircleMarkers(data = vdac.Nep, vdac.Nep$coord.Y, vdac.Nep$coord.X,
                   weight = 0.5,
                   col = "#993300", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor Nepal') %>%
  addCircleMarkers(data = vdac.SI, vdac.SI$coord.Y, vdac.SI$coord.X,
                   weight = 0.5,
                   col = "#FFDD00", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor South India 1') %>%
    addCircleMarkers(data = vdac.SRI, vdac.SRI$coord.Y, vdac.SRI$coord.X,
                   weight = 0.5,
                   col = "#669999", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor Sri Lanka 1') %>%
      addCircleMarkers(data = vdac.V1, vdac.V1$coord.Y, vdac.V1$coord.X,
                   weight = 0.5,
                   col = "#009900", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor Viet Nam V1') %>%
  addCircleMarkers(data = vdac.C4, vdac.C4$coord.Y, vdac.C4$coord.X,
                   weight = 0.5,
                   col = "#663300", 
                   radius = 5, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. destructor China C4') %>%
  
  
  
 %>%
  addLayersControl(position = "topright",
    baseGroups = c("OSM (default)", "Positron NoLabels"),
    overlayGroups = c("V. destructor K1", 
                      "V. destructor J1", 
                      "V. destructor China C1", 
                      "V. destructor China C2", 
                      "V. destructor China C3", 
                      "V. destructor China C4", 
                      "V. destructor Nepal", 
                      "V. destructor South India 1", 
                      "V. destructor Sri Lanka 1", 
                      "V. destructor Viet Nam V1",
                      "V. jacobsoni"),
    options = layersControlOptions(collapsed = TRUE))  %>%
  ## adding a title for the map
  addControl(title, position = "bottomright", className="map-title2") %>% 
  ## show the positron background prerably to the OSM layer
  showGroup("Positron NoLabels") %>% 
  hideGroup("V. jacobsoni")

```

This third map now focus on the diversity within *V. jacobsoni* on its native host *A. cerana*. Although, *V. jacobsoni* has only jumped ~10 years ago, concerns arise on its invasion potential as it could follow the steps of its sister species.

## Distribution of Varroa jacobsoni mtDNA lineages on Apis cerana
```{r jacobsoni, warning=FALSE}

vjac.Mal1 <- coord.vjac[coord.vjac$GROUP =="VJ_Malaysia_1",]
vjac.Mal2 <- coord.vjac[coord.vjac$GROUP =="VJ_Malaysia_2",]
vjac.Jav <- coord.vjac[coord.vjac$GROUP =="VJ_Java_1",]
vjac.Amb <- coord.vjac[coord.vjac$GROUP =="VJ_Ambon_1",]
vjac.Lom <- coord.vjac[coord.vjac$GROUP =="VJ_Lombok_1",]
vjac.Bal <- coord.vjac[coord.vjac$GROUP =="VJ_Bali_1",]
vjac.Beng <- coord.vjac[coord.vjac$GROUP =="VJ_Bengaluru_1",]
vjac.Sbw <- coord.vjac[coord.vjac$GROUP =="VJ_Sumbawa_1",]
vjac.Sum <- coord.vjac[coord.vjac$GROUP =="VJ_Sumatra_1",]
vjac.Flo <- coord.vjac[coord.vjac$GROUP =="VJ_Flores_1",]
vjac.Sam1 <- coord.vjac[coord.vjac$GROUP =="VJ_Samui_1",]
vjac.Sam2 <- coord.vjac[coord.vjac$GROUP =="VJ_Samui_2",]
vjac.Bor1 <- coord.vjac[coord.vjac$GROUP =="VJ_Borneo_1",]
vjac.NT1 <- coord.vjac[coord.vjac$GROUP =="VJ_North_Thailand_1",]
vjac.NT2 <- coord.vjac[coord.vjac$GROUP =="VJ_North_Thailand_2",]
vjac.NT3 <- coord.vjac[coord.vjac$GROUP =="VJ_North_Thailand_3",]
vjac.Laos <- coord.vjac[coord.vjac$GROUP =="VJ_Laos_L1",]

# Prepare the map title
tag.map.title3 <- tags$style(HTML("
  .leaflet-control.map-title3 { 
    transform: translate(-50%,20%);
    text-align: center;
    padding-left: 10px; 
    padding-right: 10px; 
    background: rgba(255,255,255,0.75);
    font-weight: bold;
    font-size: 20px;
  }
"))

title <- tags$div(
  tag.map.title3, HTML("Reported distribution of V. jacobsoni strains on A. cerana")
)  

leaflet(coord.vjac) %>% 
  addTiles(group = "OSM (default)") %>%
  ## add the layer other than default we would like to use for background
  addProviderTiles(providers$CartoDB.PositronNoLabels, group = "Positron NoLabels") %>%  
  ## adding each lineage of V. jacobsoni as a layer
  addCircleMarkers(data = vjac.Mal1, vjac.Mal1$coord.Y, vjac.Mal1$coord.X,
                   weight = 0.5,
                   col = "#4dffff", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Malaysia_1') %>%
addCircleMarkers(data = vjac.Mal2, vjac.Mal2$coord.Y, vjac.Mal2$coord.X,
                   weight = 0.5,
                   col = "#0099ff", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Malaysia_2') %>%
addCircleMarkers(data = vjac.Jav, vjac.Jav$coord.Y, vjac.Jav$coord.X,
                   weight = 0.5,
                   col = "#0000e6", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Java_1') %>%
addCircleMarkers(data = vjac.Amb, vjac.Amb$coord.Y, vjac.Amb$coord.X,
                   weight = 0.5,
                   col = "#000000", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Ambon_1') %>%
addCircleMarkers(data = vjac.Lom, vjac.Lom$coord.Y, vjac.Lom$coord.X,
                   weight = 0.5,
                   col = "#00ff00", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Lombok_1') %>%
addCircleMarkers(data = vjac.Bal, vjac.Bal$coord.Y, vjac.Bal$coord.X,
                   weight = 0.5,
                   col = "#006600", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Bali_1') %>%
addCircleMarkers(data = vjac.Beng, vjac.Beng$coord.Y, vjac.Beng$coord.X,
                   weight = 0.5,
                   col = "#006666", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Bengaluru_1') %>%
addCircleMarkers(data = vjac.Sbw, vjac.Sbw$coord.Y, vjac.Sbw$coord.X,
                   weight = 0.5,
                   col = "#ff8000", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Sumbawa_1') %>%
addCircleMarkers(data = vjac.Sum, vjac.Sum$coord.Y, vjac.Sum$coord.X,
                   weight = 0.5,
                   col = "#ffd9b3", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Sumatra_1') %>%
addCircleMarkers(data = vjac.Flo, vjac.Flo$coord.Y, vjac.Flo$coord.X,
                   weight = 0.5,
                   col = "#ffff00", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Flores_1') %>%
  addCircleMarkers(data = vjac.Sam1, vjac.Sam1$coord.Y, vjac.Sam1$coord.X,
                   weight = 0.5,
                   col = "#a64dff", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Samui_1') %>%
addCircleMarkers(data = vjac.Sam2, vjac.Sam2$coord.Y, vjac.Sam2$coord.X,
                   weight = 0.5,
                   col = "#e6ccff", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Samui_2') %>%
addCircleMarkers(data = vjac.Bor1, vjac.Bor1$coord.Y, vjac.Bor1$coord.X,
                   weight = 0.5,
                   col = "#663300", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Borneo_1') %>%
addCircleMarkers(data = vjac.NT1, vjac.NT1$coord.Y, vjac.NT1$coord.X,
                   weight = 0.5,
                   col = "#f2f2f2", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni North_Thailand_1') %>%
addCircleMarkers(data = vjac.NT2, vjac.NT2$coord.Y, vjac.NT2$coord.X,
                   weight = 0.5,
                   col = "#bfbfbf", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni North_Thailand_2') %>%
addCircleMarkers(data = vjac.NT3, vjac.NT3$coord.Y, vjac.NT3$coord.X,
                   weight = 0.5,
                   col = "#4d4d4d", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni North_Thailand_3') %>%
  addCircleMarkers(data = vjac.Laos, vjac.Laos$coord.Y, vjac.Laos$coord.X,
                   weight = 0.5,
                   col = "#cc0066", 
                   radius = 4, 
                   fillOpacity = 0.9, 
                   stroke = T, 
                   label = ~as.character(Description), 
                   group = 'V. jacobsoni Laos_1') %>%
  addLayersControl(position = "topright",
    baseGroups = c("OSM (default)", "Positron NoLabels"),
    overlayGroups = c("V. jacobsoni Malaysia_1", 
                      "V. jacobsoni Malaysia_2", 
                      "V. jacobsoni Java_1",
                      "V. jacobsoni Ambon_1", 
                      "V. jacobsoni Lombok_1",
                      "V. jacobsoni Bali_1",
                      "V. jacobsoni Bengaluru_1",  
                      "V. jacobsoni Sumbawa_1", 
                      "V. jacobsoni Sumatra_1",  
                      "V. jacobsoni Flores_1",  
                      "V. jacobsoni Samui_1", 
                      "V. jacobsoni Samui_2", 
                      "V. jacobsoni Borneo_1", 
                      "V. jacobsoni North_Thailand_1", 
                      "V. jacobsoni North_Thailand_2", 
                      "V. jacobsoni North_Thailand_3",
                      "V. jacobsoni Laos_1"),
    options = layersControlOptions(collapsed = TRUE))  %>%
  ## adding a title for the map
  addControl(title, position = "bottomright", className="map-title3") %>% 
  ## show the positron background prerably to the OSM layer
  showGroup("Positron NoLabels")
```

This interactive map aims to reach the greatest audience and to initiate participative science among researchers, bee and mites experts, beekeepers and amateur entomologists. If you are interested in improving this current database (adding on missing data and new records) or even noticed some errors, please contact *Maeva Techer*, Postdoctoral JSPS Fellow at the Okinawa Institute of Science and Technology, Japan: maeva.techer@oist.jp or maeva-angelique.techer@gmail.com


